### 2022-05-30 paper.Rmd
discussion

In this paper, I presented the k-plus criterion for anticlustering. K-plus anticlustering focuses on dividing a data set into subgroups in such a way that these groups are similar to each other. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. The k-plus family may be used to equalize the variance as well, any higher order moments such as skewness and kurtosis, as well as covariances. Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data. A simulation study and examples on real norming data indicated that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.

This work indicates that k-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups, given that similarity in means and variances can usually be achieved succesfully at the same time. Depending on a user's needs, equalizing higher order moments can also be requested, but it should be noted that fulfilling several criteria at the same time may be difficult to achieve for small data sets. In any way, a user can just try out different objectives, i.e., start greedy by requesting that multiple distribution moments are equalized, and if the results for the means and standard deviations are subpar, try out fewer criteria. K-plus anticlustering was also shown to outperform diversity anticlustering with regard to the specific objectives that are included in the optimization process; diversity is still a reasonable all-arounder that strives for overall betwee-group similarity. Which method should be employed depends on a user's preferences. In doubt, several anticlustering methods can be applied sequentially and the results can be compared directly. Unlike in statistical methods, there is no problem with trying out different approaches when it comes to anticlustering.

@@

- This paper focused on the presentation of new objective function, the bicriterion optimization process might be improved [e.g., by using a direct algorithm as @brusco2019 did]. From the algorithmic point of view, this paper did not strive to apply the latest cutting edge methodology. For most practical purposes, however, I assume that optimizing the k-plus criterion will yield very satisfactory results for researchers. I have already worked with researchers planning their experiments who are generally amazed by the groupd that are returned by k-plus anticlustering.

Moreover, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it. Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes. It was of particular interest to compare the k-plus objective with the popular diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups, whereas the diversity maintains an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion regains a lot of attractivity: when optimizing for similarity in means and variance, k-plus anticlustering clearly outperformed the diversity criterion in the simulation study. It should also be noted that the k-plus criterion should almost always be preferred to standard k-means. This is because even for low $N$, k-plus equalizes means quite well, and for larger $N$ the practical differences between k-means and k-plus vanish completely. A reasonable exception is given when the data set consists of binary variables where mean and standard deviation are directly related.

@@

## Future research

- Future research: Application of more sophisticated bicriterion optimization algorithms are interesting

- extending k-means anticlustering on maximizing similarity with regard to covariances? Mahalanobis distance?

# General thoughts on anticlustering

@@

>>>>

In this paper, I presented the k-plus criterion for anticlustering. K-plus anticlustering focuses on dividing a data set into subgroups in such a way that these groups are similar to each other. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. The k-plus criterion however may be used to also equalize the variance between groups, and any higher order moments such as skewness and kurtosis, as even covariances. Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data.

A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria. In particular, this work indicates that k-plus anticlustering should be a replacement of the classical k-means approach when striving for between-group similarity. K-means anticlustering disregards any distribution characteristics other than the mean; k-plus anticlustering ensures between-group similarity with regard to multiple criteria while still taking into account differences in means. For large data sets, there is virtually no reason to use k-means anticlustering because the practical differences between k-means and k-plus regarding similarity in means vanish completely. A reasonable exception is given when the input data consists of binary variables because in this case the mean and the standard deviation are directly related.

The results showed that fulfilling several criteria at the same time may be difficult to achieve for small data sets. Still, users can try out different objectives and see which specification fulfils their needs. A reasonable approach it to start greedy, i.e., requesting multiple distribution moments to be equalized, and only if the results are subpar, go for fewer criteria. Unlike in statistical analysis, there is no problem with "just trying out" different approaches when it comes to anticlustering.

K-plus anticlustering was also shown to outperform diversity anticlustering with regard to the specific objectives that are included in the optimization process; diversity is still a reasonable all-arounder that strives for overall betwee-group similarity. Which method should be employed depends on a user's preferences. In doubt, several anticlustering methods can be applied sequentially and the results can be compared directly.

@@

This paper focused on the presentation of new objective function but the did not properly consider the multicriterion nature of the k-plus objective. Instead, a simple (unweighted) sum approach was applied to optimize the k-plus objective, independent of the number of criteria that were incorporated. Other approaches, such as direct algorithms, may yield improved results and better possibilities to select partitionings according to differently weighted criteria. @brusco2019 showed how direct algorithms can be used to optimize multiple critera in anticlustering applications. Their study focused on an approach to find the pareto set of solutions that simultaneously optimized diversity and dispersion, thereby maximizing within-group heterogeneity with regard to both criteria.[^inanticlust]

[^inanticlust]: The `anticlust` package now also includes an implementation of the bicriterion anticlustering algorithm by @brusco2019.

Focusing on the objetive function only and not on specialized algorithmic optimization approaches followed two reasonings. First, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it. Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes. It was of particular interest to compare the k-plus objective with the popular diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups, whereas the diversity maintains an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion regains a lot of attractivity: when optimizing for similarity in means and variance, k-plus anticlustering clearly outperformed the diversity criterion in the simulation study. Second, the applications show that the simple sum approach using a general local maximum search yields very satisfying results in practice. It seems that the space of solutions in anticlustering settings includes partitionings that fulfil multiple objectives and these partitionings can be found using a local maximum search.

@@

## Closing thoughts on the nature of anticlustering

@@

## Conclusion

This paper introduced k-plus anticlustering for maximizing between-group similarity, improving the traditional k-means approach. K-plus anticlustering can be adopted whenever a data set has to be partitioned into equivalent parts. The method is accessible and easily usable via the R package `anticlust`.
